|

### 2015-12-09 manuscript/manuscript_pve.Rmd
failed changes to doc.x
this manuscrupt saved as thesis draft 1

Interpreting limitations to plant growth requires understanding of the balance between carbon (C) source and sink activity. This study used manipulations of soil volume to test how growth is coupled to physiology, C allocation, and sink activity in *Eucalyptus tereticornis* seedlings. We grew individual seedlings in a large range of container sizes and planted containers flush to the soil alongside naturally sown (free) seedlings. We developed a seedling growth model that utilized leaf photosynthesis rates (*A~n~*) to allocate daily C uptake towards mass growth of stems, leaves and roots. Reduced soil volume was expected to induce rapid negative effects on growth and physiology compared to free seedlings. It was hypothesized that the soil volume effect would be largest in the smallest containers, negatively impacting mass partitioning belowground. An accumulation of leaf non-structural carbohydrates, resulting from reduced belowground sink strength, was expected to correlate to reductions in photosynthetic capacity. We observed a negative effect of container volume on aboveground growth soon after the experiment started. Although growth was consistently different across soil volumes, dry mass partitioning to leaves, stems and roots was unchanged after 120 days. Photosynthetic capacity was significantly reduced in containers compared to free seedlings, and was related to both leaf nitrogen content and starch accumulation. We then asked whether the observed reductions in *A~n~* explained the observed differences in seedling biomass. We found that although belowground sink limitation resulted in down regulation of *A~n~*, these reductions were not large enough to explain observed growth responses. Thus, as *A~n~* and growth were not tightly coordinated, excess photosynthetic C not attributed to biomass resulted in seedlings with soil volume restriction. These results highlight the need to further utilize mass balance approaches when evaluating plant C allocation and confirms that important feedbacks exist between belowground sink strength and leaf C uptake.

@@

##2.4.1 Growth and morphology
Plant height, diameter and leaf area diverged between container volumes soon after start of the experiment (Figure `r allomfig`). First, seedling leaf area significantly diverged between soil volumes (P < 0.029) during the 5^th^ week of the experiment. Following this period both height (8^th^) week) and then diameter (9^th^ week) significantly deviated across soil volumes (P < 0.045 & 0.035, respectively). The large reductions in height gain and total leaf area in smaller compared to larger containers continued throughout the experiment. In this field study, colder temperatures and reductions in total PPFD per day (Figure `r airvarsfig`) likely slowed the growth of seedlings in the final weeks of the experiment (Figure `r allomfig`). Seedlings maintained diameter growth throughout the experiment, although marginal with smaller soil volumes in the final month. Final seedling height significantly increased with increasing soil volume (P < 0.001). Increases in both final stem diameter (P < 0.001) and cumulative leaf area (both P < 0.001) were found with increasing soil volume and these differences were driven mainly by the largest container and the free seedling treatments.
\
```{r allometry, fig.width=6, fig.height=7.8, echo=FALSE,  dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/allometryplotting.R")
```  
\
**Figure 2.1**. Soil volume treatment means &pm; standard error of height growth (a), diameter growth (b), and interpolated seedling leaf area (c) measured weekly of *Eucalyptus tereticornis* seedlings across the experiment duration in 2013.
\
```{r airvars, fig.width=6, fig.height=7.8, echo=FALSE, dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/airvars_plotting.R")
```  
\
**Figure 2.2**. Daily maximum and minimum temperature (a), total daily PPFD (b), and daily maximum vapour pressure deficit (c) across the experiment duration in 2013.

>>>>

Interpreting limitations to plant growth requires understanding of the balance between carbon (C) source and sink activity. This study used manipulations of soil volume to test how growth is coupled to physiology, C allocation, and sink activity in *Eucalyptus tereticornis* seedlings. We grew individual seedlings in a large range of container sizes and planted containers flush to the soil alongside naturally sown (free) seedlings. We developed a seedling growth model that utilized leaf photosynthesis rates (*A~n~*) to allocate daily C uptake towards mass growth of stems, leaves and roots. Reduced soil volume was expected to induce rapid negative effects on growth and physiology compared to free seedlings. It was hypothesized that the soil volume effect would be largest in the smallest containers, negatively impacting mass partitioning belowground. An accumulation of leaf non-structural carbohydrates, resulting from reduced belowground sink strength, was expected to correlate to reductions in photosynthetic capacity. We observed a negative effect of container volume on aboveground growth soon after the experiment started. Although growth was consistently different across soil volumes, dry mass partitioning to leaves, stems and roots was unchanged after 120 days. Photosynthetic capacity was significantly reduced in containers compared to free seedlings, and was related to both leaf nitrogen content and starch accumulation. We then asked whether the observed reductions in *A~n~* explained the observed differences in seedling biomass. We found that although belowground sink limitation resulted in down regulation of *A~n~*, these reductions were not large enough to explain observed growth responses. Thus, as *A~n~* and growth were not tightly coordinated, excess photosynthetic C not attributed to biomass resulted in seedlings with soil volume restriction. These results highlight the need to further utilize mass balance approaches when evaluating plant C allocation and confirm that important feedbacks exist between belowground sink strength and leaf C uptake.

@@

##2.4.1 Growth and morphology
Plant height, diameter and leaf area diverged between container volumes soon after start of the experiment (Figure `r allomfig`). First, seedling leaf area significantly diverged between soil volumes (P < 0.029) during the 5^th^ week of the experiment. Following this period both height (8^th^) week) and then diameter (9^th^ week) significantly deviated across soil volumes (P < 0.045 & 0.035, respectively). The large reductions in height gain and total leaf area in smaller compared to larger containers continued throughout the experiment. In this field study, colder temperatures and reductions in total PPFD per day (Figure `r airvarsfig`) likely slowed the growth of seedlings in the final weeks of the experiment (Figure `r allomfig`). Seedlings maintained diameter growth throughout the experiment, although marginal with smaller soil volumes in the final month. Final seedling height significantly increased with increasing soil volume (P < 0.001). Increases in both final stem diameter (P < 0.001) and cumulative leaf area (both P < 0.001) were found with increasing soil volume and these differences were driven mainly by the largest container and the free seedling treatments.
\
```{r allometry, fig.width=6, fig.height=7.6, echo=FALSE,  dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/allometryplotting.R")
```  
\
**Figure 2.1**. Soil volume treatment means &pm; standard error of height growth (a), diameter growth (b), and interpolated seedling leaf area (c) measured weekly of *Eucalyptus tereticornis* seedlings across the experiment duration in 2013.
\
```{r airvars, fig.width=6, fig.height=7.6, echo=FALSE, dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/airvars_plotting.R")
```  
\
**Figure 2.2**. Daily maximum and minimum temperature (a), total daily PPFD (b), and daily maximum vapour pressure deficit (c) across the experiment duration in 2013.
